---
title: "STAT400 - Homework 2"
author: "Your Name"
date: "Due 9/17/2019 by 4pm"
output: pdf_document
---

```{r, message=FALSE, warning=FALSE}
## Some useful packages
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
```

The City of Fort Collins maintains a web portal to host data sets that are available to the public in a format that citizens can easily search, filter, and utilize (https://www.fcgov.com/opendata/). In particular, they have released the earnings of city employees from 2010-2018.

Each of these data sets are included in this repo under the `data/` folder.

1. Write a function that will take a year (2010-2018) as an input parameter and do the following:
    - Load the corresponding CSV into `R` (i.e. "data/[year]_Fort_Collins_City_Employee_Earnings.csv", see `paste()`).
    - Extract the `Job Title`, `Department` and `Earnings` columns.
    - Change the column names to `job_title`, `department`, and `earnings`.
    - Make `job_title` and `department` values lower case (see `tolower()` and `mutate_at()`).
    - Add a column called `year` with the appropriate year.
    
    ```{r}
    ## Write your function here
    load_and_format <- function(year) {
      
    }
    ```
   
2. Apply your function to all years available of the city employees data (2010-2018) and store your results in a single data frame (see `rbind` for binding rows of data frames together).

    ```{r}
    ## Apply your function to the years available
    years <- 2010:2018
    
    ```

3. For those departments that have over 25 unique job titles, make a plot that depicts earnings over time by department in the city of Fort Collins.

    ```{r}
    ## Visualize your data
    
    ```
