---
title: "STAT400 - Homework 6"
author: "Your Name"
date: "Due 10/15/2020 by 4pm"
output: pdf_document
---
Be sure to `set.seed(400)` at the beginning of your homework.

```{r}
#reproducibility
set.seed(400)
```

```{r, message=FALSE}
# useful libraries
library(tidyverse)
```


1. Develop two Monte Carlo integration approaches to estimate $\int\limits_0^5 x^2 \exp(-x) dx$. (You must use different distributions in the two approaches). Check your answer using the `integrate()` function.

    ```{r}
    # method 1
    
    # method 2
    
    # check answers with integrate()
    
    ```

2. Estimating the cdf of a normal distribution. Use $m = 1000$.
    
    a. Implement all 3 methods that we discussed in class (Example 1.7, Page 9-10 of Ch. 6 Notes) to estimate the cdf of a normal distribution $\Phi(x)$. Note that you will need to show some derivations for method 2.
    b. Compare your estimates with the output from the `R` function `pnorm()` for $x = 0.5, 1, 2, 3$. Summarise your findings comparing the performance of the methods.
    c. For each method, compute an estimate of the variance of your Monte Carlo estimate of $\Phi(2)$. Summarise your findings.
    d. For each method, compute a $95\%$ confidence interval for $\Phi(2)$. Summarise your findings. Which CI is narrower and why does that matter?
    
    ```{r}
    # method 1
    
    # method 2
    
    # method 3
    
    # compare to pnorm for x = 0.5, 1, 2, 3
    
    # compute estimates of variance for Phi(2)
    
    # compute CIs for Phi(2)
    
    ```
    
3. Compute a Monte Carlo estimate $\hat{\theta}_1$ of
    $$
    \theta = \int\limits_0^{0.5} e^{-x} dx
    $$
    by sampling from the Uniform($0, 0.5$) and estimate the variance of $\hat{\theta}_1$. Find another Monte Carlo estimator $\hat{\theta}_2$ by sampling from the Exponential(1) distribution and estimating its variance.
    
    Which of the variances (of $\hat{\theta}_1$ or $\hat{\theta}_2$) is smaller?
    
    ```{r}
    # number of samples
    m <- 1000
    
    ## first MC estimator and variance
    ## use the Unif(0, .5) dsn
    
    ## second MC estimator and variance
    ## use the Exp(1) dsn
    
    ## Compare estimated variances
    
    ```
    
    
