---
title: "STAT400 - Homework 9"
author: "Your Name"
date: "Due 11/10/2022 by 4pm"
output: pdf_document
---

Be sure to `set.seed(400)` at the beginning of your homework.

```{r}
#reproducibility
set.seed(400)
```

```{r, message=FALSE}
# useful libraries
library(tidyverse)
```

1. Use the Monte Carlo simulation to investigate whether the empirical Type I error rate of the $t$-test is approximately equal to the nominal significance level when the sampled population is non-normal.

    a. For $n = 5, 10, 30, 100, 500, 1000$, investigate the empirical type I error for a test of $H_0: \mu = 1$ vs. $H_a: \mu \not= 1$ when $X_1, \dots, X_n \sim \chi^2(1)$ with $m = 2000$ Monte Carlo samples with nominal $\alpha = .05$.
    
    b. For $n = 5, 10, 30, 100, 500, 1000$, investigate the empirical type I error for a test of $H_0: \mu = 1$ vs. $H_a: \mu \not= 1$ when $X_1, \dots, X_n \sim Unif[0, 2]$ with $m = 2000$ Monte Carlo samples with nominal $\alpha = .05$.
    
    c. For $n = 5, 10, 30, 100, 500, 1000$, investigate the empirical type I error for a test of $H_0: \mu = 1$ vs. $H_a: \mu \not= 1$ when $X_1, \dots, X_n \sim Exponential(1)$ with $m = 2000$ Monte Carlo samples with nominal $\alpha = .05$.
    
    d. Compare your results in a.-c. in a table. What can you say about the departures from Normality as they relate to the Type I error rate of the $t$-test?
    
    ```{r}
    # function to compute empirical alpha based on n, alpha, m, and the generating distn
    
    # a. \chi^2(1)
    
    # b. Unif[0, 2]
    
    # c. Exp(1)
    
    # make table to compare results
    
    ```

