---
title: "STAT400 - Homework 10"
author: "Your Name"
date: "Due 12/03/2020 by 4pm"
output: pdf_document
---

Be sure to `set.seed(400)` at the beginning of your homework. Please use $2000$ bootstrap samples in the problems below.

```{r}
#reproducibility
set.seed(400)

# bootstrap samples
B <- 2000
```

```{r, message=FALSE}
# useful libraries
library(tidyverse)
```

1. Nike has hired you to help analyze their data on their customers who run. They want to make sure that you understand how their running gear fits their customers. A sample of 25 randomly selected customers was selected, and the customers were asked to submit their weights. The data:

    ```{r}
      wt <- c(149, 136, 139, 117, 137, 132, 122, 130, 134, 153, 140, 151, 203, 143, 145, 123, 127, 146, 139, 118, 143, 123, 133, 153, 136)
    ```

    a. Calculate the sample standard deviation $s$ for these weights.
    b. To do the following, use the `boot` and `simpleboot` packages as shown in the class handouts.
        i. Compute the bootstrap bias and standard error for $s$.
        ii. Plot a histogram and qq-plot of the bootstrap distribution.
        iii. Based on these results: (1) Is there evidence of bias and skewness of the bootstrap distribution for $s$? (2) Is it appropriate to assume that the distribution of $s$ is normally distributed?
        iv. Construct 4 types of intervals that we discussed in class by using `type=c("norm","basic", "perc", "bca")` in the boot.ci command
        v. Plot the four intervals onto a histogram of the sampling distribution using the command `geom_segment`.
    c. Construct a "studentized" bootstrap $t$ CI and also plot it onto your histogram of the sampling distribution.
    d. What final result would you report to Nike? Explain your reasoning.
    
2. For the `Verizon` dataset from the class handouts construct a 95% CI for the difference of two medians. Construct 4 types of intervals that we discussed in class by using `type=c("norm","basic", "perc", "bca")` in the boot.ci command.


    a. Are the intervals similar for all the methods? Why or why not?
    b. Let $\tilde{\mu}_1$ = the population median repair time for ILEC customers and $\tilde{\mu}_2$ = the population median of repair time for CLEC customers. Based on the results of the BCa interval, would you reject this hypothesis? Explain your answer.
        $$
        H_0: \tilde{\mu}_1 = \tilde{\mu}_2 \\
        H_a: \tilde{\mu}_1 \not= \tilde{\mu}_2 \\
        $$
        
3. This data set is the `cars` data in `R`. The goal is to create a regression model about the relationship between stopping distance (`dist`) and speed (`speed`) in cars.

    a. Create a scatter plot of `speed` vs. `dist` in `mpg` and describe the relationship.
    b. Fit a simple linear model of `dist` on `speed`. Describe the result, including diagnostic plots, and create 95% CIs for the coefficients.
    c. Perform the paired bootstrap and compute the bootstrap bias and standard error for the coefficients. Construct 4 types of intervals that we discussed in class by using `type=c("norm","basic", "perc", "bca")` in the boot.ci command.
    d. Perform the bootstrap using the residuals and compute the bootstrap bias and standard error for the coefficients. Construct 4 types of intervals that we discussed in class by using `type=c("norm","basic", "perc", "bca")` in the boot.ci command.
    e. Which method (simple linear regression, paired bootstrap, or bootstrap using the residuals) would you use for this dataset and why?