---
title: "STAT400 - Homework 9"
author: "Your Name"
output: pdf_document
---

Be sure to `set.seed(400)` at the beginning of your homework.

```{r}
#reproducibility
set.seed(400)
```

```{r, message=FALSE}
# useful libraries
library(tidyverse)
```

1. Use the Monte Carlo simulation to investigate whether the empirical Type I error rate of the $t$-test is approximately equal to the nominal significance level when the sampled population is non-normal.

    a. For $n = 5, 10, 30, 100, 500, 1000$, investigate the empirical type I error for a test of $H_0: \mu = 1$ vs. $H_a: \mu \not= 1$ when $X_1, \dots, X_n \sim \chi^2(1)$ with $m = 2000$ Monte Carlo samples with nominal $\alpha = .05$.
    
    b. For $n = 5, 10, 30, 100, 500, 1000$, investigate the empirical type I error for a test of $H_0: \mu = 1$ vs. $H_a: \mu \not= 1$ when $X_1, \dots, X_n \sim Unif[0, 2]$ with $m = 2000$ Monte Carlo samples with nominal $\alpha = .05$.
    
    c. For $n = 5, 10, 30, 100, 500, 1000$, investigate the empirical type I error for a test of $H_0: \mu = 1$ vs. $H_a: \mu \not= 1$ when $X_1, \dots, X_n \sim Exponential(1)$ with $m = 2000$ Monte Carlo samples with nominal $\alpha = .05$.
    
    d. Compare your results in a.-c. in a table. What can you say about the departures from Normality as they relate to the Type I error rate of the $t$-test?
    
    ```{r}
    # function to compute empirical alpha based on n, alpha, m, and the generating distn
    
    # a. \chi^2(1)
    
    # b. Unif[0, 2]
    
    # c. Exp(1)
    
    # make table to compare results
    
    ```
    
2. Suppose $X_1, \dots, X_n$ is a random sample from a $N(\mu, \sigma^2)$ distribution. Consider the test $H_0: \mu = 500$ vs. $H_a = \mu \not= 500$ with $\alpha = 0.05$. Then under the alternative hypothesis,
    $$
    T^* = \frac{\overline{X} - 500}{s/{\sqrt{n}}} \sim t_{n = 1}.
    $$
    for $\mu_a \not= 500$. We will examine the impact of effect size on power. Use $m = 1000$ for the number of Monte Carlo replications and $\sigma = 100$.
    
    a. Set $n = 20$. Plot the estimated power or $\mu_a$ ranging between $350$ to $650$.
    b. Set $n = 100$ and repeat part a. Plot both power curves on the same plot. What happens to power as $n$ increases?
    c. What is the estimated power when $\mu_a = 450$ for $n = 20$ and $n = 100$?
    d. When $n = 20$, what effect size will you need to achieve power of approximately $0.80$?
    
    ```{r}
    ## write function to estimate power via monte carlo for different values of n and mu_a
    
    ## make plots
    
    ```

