---
title: "STAT400 - Homework 7"
author: "Your Name"
date: "Due 10/18/2022 by 4pm"
output: pdf_document
---
Be sure to `set.seed(400)` at the beginning of your homework.

```{r}
#reproducibility
set.seed(400)
```

```{r, message=FALSE}
# useful libraries
library(tidyverse)
```
 
1. Find two importance functions $\phi_1$ and $\phi_2$ that are supported on $(1, \infty)$ and are "close" to
    $$
    h(x) = \frac{x^2}{\sqrt{2\pi}} e^{-x^2/2}, \qquad x > 1.
    $$
    
    Which of your two importance functions should produce the smallest variance in estimating
    $$
    \int\limits_1^\infty \frac{x^2}{\sqrt{2\pi}} e^{-x^2/2} dx
    $$
    by importance sampling? Explain. 
    
    **Hint:** You will need to create plots of $\phi_1$, $\phi_2$, and $g(x)f(x)$ as well as $g(x)f(x)/\phi(x)$ to answer this question.
    
    ```{r}
    ## create functions for h(x), phi_1(x), and phi_2(x)
    
    ## plot h(x), phi_1(x), and phi_2(x) together
    
    ## plot h(x)/phi_1(x) and h(x)/phi_1(x)
    
    ## which is better? Why?
    ```
    
2. Obtain a Monte Carlo estimate of     
    $$
    \int\limits_1^\infty \frac{x^2}{\sqrt{2\pi}} e^{-x^2/2} dx
    $$
    using importance sampling with the two importance sampling functions you chose in Problem 4 ($\phi_1$ and $\phi_2$). Obtain an estimate of the variance for each and compare.
    
    ```{r}
    ## number of samples to use
    m <- 100000
    
    ## estimate the integral using importance sampling
    ## phi 1
    
    
    ## phi 2
    
    
    ## estimates
    ## true value for theta is approx 0.4006
    
    
    ## compare variances
    
    ```